Using Monte Carlo Algorithms to Achieve Resiliency \& Performance at Scale for Linear and Nonlinear Solver Applications

### 2016-01-26 doc/ecp/pres.tex
Table formatting.
Almost finished.

  \begin{table}[!h]
  \centering
  \begin{tabular}{|c|c|c|c|}
  \hline
  \textbf{Block size} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$
  &$\rho(\hat{H})$\\
  \hline
   3 & 1.4352 & $0.9804$ & $0.9531$\\
  \hline
   9 & 3.0220 & $0.9790$ & $0.9674$\\
  \hline
   33 & 9.8164 & $0.9783$  & $0.9774$\\
  \hline
  \end{tabular}
  \caption{$\rho(H)$ and $\rho(\hat{H})$ for advection--diffusion
  matrix with block diagonal preconditioning.}
  \label{tab:ifiss_block}
  \end{table}
\end{frame}

@@

  \begin{table}[!h]
  \centering
  \begin{tabular}{|c|c|c|c|}
  \hline
  \textbf{Drop tolerance} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$
  &$\rho(\hat{H})$\\
  \hline
   0.05 & 8.2797 & $0.9610$ & $1.0231$\\
  \hline
   0.01 & 33.1390 & $0.8668$ & $0.8279$\\
  \hline
  \end{tabular}
  \caption{$\rho(H)$ and $\rho(\hat{H})$ for 2D
  Laplacian with AINV preconditioning.}
  \label{tab:lap2d_ainv}
  \end{table}
\end{frame}

@@

  \begin{table}[h]
    \begin{center}
      \caption{2D parabolic problem with AINV preconditioning.}
      \begin{tabular}{|c|c|c|c|}\hline
        Node & Iterations & Time (sec) & Relative Residual \\\hline
        CPU & 6 & 992.23 & $1.23\times 10^{-7}$ \\\hline
        GPU & 6 & 29.83 & $1.53\times 10^{-7}$ \\\hline
      \end{tabular}
    \end{center}
  \end{table}

@@

      \small{
        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}&
              \multicolumn{1}{c}{NK}&
              \multicolumn{1}{c}{FANM}&
              \multicolumn{1}{c}{NR}\\
              \hline
              Ra=\sn{1}{3} & 5 & 5 & 5\\
              Ra=\sn{1}{4} & 7 & 7 & 7\\
              Ra=\sn{1}{5} & 9 & 10 & 9\\
              Ra=\sn{1}{6} & 11 & 11 & 11\\
              %%
              \hline\hline
            \end{tabular}`
          \end{center}
          \caption{Nonlinear iterations.}
        \end{table}

        \begin{table}[h!]
          \begin{center}
            \begin{tabular}{lccc}\hline\hline
              \multicolumn{1}{l}{Benchmark}&
              \multicolumn{1}{c}{GMRES}&
              \multicolumn{1}{c}{MCSA}&
              \multicolumn{1}{c}{Richardson}\\
              \hline
              Ra=\sn{1}{3} & 32 & 18 & 38\\
              Ra=\sn{1}{4} & 23 & 17 & 34\\
              Ra=\sn{1}{5} & 25 & 20 & 34\\
              Ra=\sn{1}{6} & 39 & 25 & 48\\
              %%
              \hline\hline
            \end{tabular}
          \end{center}
          \caption{Total linear solver iterations.}
        \end{table}
      }

@@

%%---------------------------------------------------------------------------%%
\subsection{ECP R\&D}
%%---------------------------------------------------------------------------%%

\begin{frame}{ECP Coordinated Reasearch}

@@

  MC methods are used in many DOE application areas:
  \begin{itemize}
    \setlength{\itemsep}{-0.2\baselineskip}
  \item high-energy density physics
  \item astrophysics
  \item nuclear engineering
  \item particle accelerators and neutron spallation source
  \item quantum chemistry
  \item visualization
  \item fluid dynamics
  \item combustion
  \item computational biology
  \end{itemize}

@@

\begin{frame}{ECP R\&D I: Time-dependent Problems}

  Investigate use in time-dependent applications, e.g. non-linear radiation
  hydrodynamics for high-energy density physics applications:
  \begin{itemize}
  \item Used as inner linear solver in a non-linear (Newton) iteration
  \item Time-dependent applications generally have strong diagonal-dominance
    that makes MCSA competitive with Krylov methods
  \item MCSA has resiliency and parallel efficiency benefits over conventional
    subspace methods
  \end{itemize}

@@

\begin{frame}{ECP R\&D II: Monte Carlo Particle Transport}

  MC particle transport has applications in many areas across the DOE complex
  \begin{itemize}
    \setlength{\itemsep}{-0.2\baselineskip}
  \item high-energy density physics
  \item neutron spallation source
  \item nuclear engineering
  \item astrophysics
  \item combustion
  \end{itemize}

@@

  There are significant challenges in porting random walk algorithms to modern
  heterogeneous computing devices
  \begin{enumerate}
  \item Random walks are dominated by branching statements $\rightarrow$
    thread divergence
  \item Runtime is often dominated by sampling and interpolating large fields
    of data (keeping the GPUs fed)
  \end{enumerate}

@@

\begin{frame}{ECP R\&D III: Particle Methods}

>>>>

  \begin{table}
  \centering
  \begin{tabular}{cccc}
  \toprule
  \textbf{Block size} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$
  &$\rho(\hat{H})$\\
  \midrule
   3 & 1.4352 & $0.9804$ & $0.9531$\\
   9 & 3.0220 & $0.9790$ & $0.9674$\\
   33 & 9.8164 & $0.9783$  & $0.9774$\\
  \bottomrule
  \end{tabular}
  \caption{$\rho(H)$ and $\rho(\hat{H})$ for advection--diffusion
  matrix with block diagonal preconditioning.}
  \label{tab:ifiss_block}
  \end{table}
\end{frame}

@@

  \begin{table}
  \centering
  \begin{tabular}{cccc}
  \toprule
  \textbf{Drop tolerance} & \textbf{$\frac{nnz(H)}{nnz(A)}$}& $\rho(H)$
  &$\rho(\hat{H})$\\
  \midrule
   0.05 & 8.2797 & $0.9610$ & $1.0231$  \\
   0.01 & 33.1390 & $0.8668$ & $0.8279$ \\
  \bottomrule
  \end{tabular}
  \caption{$\rho(H)$ and $\rho(\hat{H})$ for 2D
  Laplacian with AINV preconditioning.}
  \label{tab:lap2d_ainv}
  \end{table}
\end{frame}

@@

  \begin{table}[h]
    \begin{center}
      \caption{2D parabolic problem with AINV preconditioning.}
      \begin{tabular}{cccc}
        \toprule
        Node & Iterations & Time (sec) & Relative Residual \\
        \midrule
        CPU & 6 & 992.23 & $1.23\times 10^{-7}$ \\
        GPU & 6 & 29.83 & $1.53\times 10^{-7}$ \\
        \bottomrule
      \end{tabular}
    \end{center}
  \end{table}

@@

      \small{
        \begin{table}
          \begin{center}
            \begin{tabular}{lccc}
              \toprule
              \multicolumn{1}{l}{Benchmark}&
              \multicolumn{1}{c}{NK}&
              \multicolumn{1}{c}{FANM}&
              \multicolumn{1}{c}{NR}\\
              \midrule
              Ra=\sn{1}{3} & 5 & 5 & 5\\
              Ra=\sn{1}{4} & 7 & 7 & 7\\
              Ra=\sn{1}{5} & 9 & 10 & 9\\
              Ra=\sn{1}{6} & 11 & 11 & 11\\
              %%
              \bottomrule
            \end{tabular}`
          \end{center}
          \caption{Nonlinear iterations.}
        \end{table}

        \begin{table}
          \begin{center}
            \begin{tabular}{lccc}
              \toprule
              \multicolumn{1}{l}{Benchmark}&
              \multicolumn{1}{c}{GMRES}&
              \multicolumn{1}{c}{MCSA}&
              \multicolumn{1}{c}{Richardson}\\
              \midrule
              Ra=\sn{1}{3} & 32 & 18 & 38\\
              Ra=\sn{1}{4} & 23 & 17 & 34\\
              Ra=\sn{1}{5} & 25 & 20 & 34\\
              Ra=\sn{1}{6} & 39 & 25 & 48\\
              %%
              \bottomrule
            \end{tabular}
          \end{center}
          \caption{Total linear solver iterations.}
        \end{table}
      }

@@

%%---------------------------------------------------------------------------%%
\subsection{R\&D in ECP}
%%---------------------------------------------------------------------------%%

\begin{frame}{Monte Carlo R\&D in ECP}

@@

  MC methods are used in many DOE application areas:
  \begin{itemize}
    \setlength{\itemsep}{-0.2\baselineskip}
  \item fluid dynamics
  \item combustion
  \item quantum chemistry
  \item visualization
  \item nuclear engineering
  \item high-energy density physics
  \item astrophysics
  \item particle accelerators and neutron spallation source
  \item computational biology
  \end{itemize}

@@

\begin{frame}{Monte Carlo for Time-dependent Applications in ECP}

  Investigate use in time-dependent applications, e.g. fluid flow, compustion,
  non-linear radiation hydrodynamics for high-energy density physics
  applications:
  \vfill
  \begin{itemize}
  \item Used as inner linear solver in a non-linear (Newton) iteration
  \vfill
  \item Time-dependent applications generally have strong diagonal-dominance
    that makes MCSA competitive with Krylov methods
  \vfill
  \item MCSA has resiliency and parallel efficiency benefits over conventional
    subspace methods
  \end{itemize}

@@

\begin{frame}{Monte Carlo for Particle Transport Applications in ECP}

  MC particle transport has applications across the DOE complex
  \begin{itemize}
    \setlength{\itemsep}{-0.2\baselineskip}
  \item combustion
  \item nuclear engineering
  \item rarefied gas dynamics
  \item astrophysics
  \item high-energy density physics
  \end{itemize}

@@

  There are significant challenges in porting random walk algorithms to modern
  heterogeneous computing devices
  \begin{enumerate}
    \setlength{\itemsep}{-0.2\baselineskip}
  \item Random walks are dominated by branching statements $\rightarrow$
    thread divergence
  \item Runtime is often dominated by sampling and interpolating large fields
    of data (keeping the GPUs fed)
  \end{enumerate}

@@

\begin{frame}{General Particle Applications in ECP}
